---
title: "Analysis: Stroop pilot"
author: "Your name"   # <- put your name here
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE)
library(tidyverse)
library(jsonlite)
```

This notebook reads the CSVs written by `npm run export -- --experiment stroop-pilot-a`
and runs the planned analyses for the Stroop task. Set `experiment_id` in the first
code chunk below to match `EXPERIMENT.id` in `experiment.js`, and put your name in the
`author:` line at the top of this file. Replace the sections below with
the confirmatory analysis you preregistered; keep exploratory analyses clearly separated.

## Load data

`trials.csv` is long format: one row per jsPsych trial, one participant per `participant_id`.
Columns that hold objects (for example the survey plugin's `response`) are JSON strings.

```{r load}
experiment_id <- "stroop-pilot-a"
raw_dir <- file.path("..", "data", "raw", experiment_id)

participants <- read_csv(file.path(raw_dir, "participants.csv"), show_col_types = FALSE)
trials <- read_csv(file.path(raw_dir, "trials.csv"), show_col_types = FALSE)
errors <- read_csv(file.path(raw_dir, "errors.csv"), show_col_types = FALSE)

glimpse(participants)
```

## Exclusions

Document every exclusion. Record how many participants each rule removes.

```{r exclusions}
n_started <- nrow(participants)
complete <- participants |> filter(completed)

# Test trials only. Practice trials (task "stroop_practice") are never analyzed.
stroop_all <- trials |>
  filter(task == "stroop") |>
  mutate(correct = as.logical(correct), timed_out = as.logical(timed_out), rt = as.numeric(rt))

# Example performance criterion: at least 75% of test trials correct (timeouts count as
# errors). Replace this with the criterion you preregistered.
stroop_accuracy <- stroop_all |>
  group_by(participant_id) |>
  summarise(
    acc = mean(correct, na.rm = TRUE),
    all_timed_out = all(timed_out),
    .groups = "drop"
  )

# A participant who never answered a single speeded trial did not really take part: this
# is what a phone or tablet produces, and what an abandoned-but-left-open tab produces.
keep <- complete |>
  left_join(stroop_accuracy, by = "participant_id") |>
  filter(!is.na(acc), !all_timed_out, acc >= 0.75)

# Exclusion funnel. `.drop = FALSE` on a factor keeps zero rows visible: a step that
# removes everyone must show as 0, not vanish from the table.
funnel_levels <- c("started", "completed", "passed accuracy criterion")
bind_rows(
  participants |> mutate(step = "started"),
  complete |> mutate(step = "completed"),
  keep |> mutate(step = "passed accuracy criterion")
) |>
  mutate(step = factor(step, levels = funnel_levels)) |>
  count(step, .drop = FALSE)

message("Participants: ", n_started, " started / ", nrow(complete), " completed / ",
        nrow(keep), " kept for analysis.")

# Participants excluded because they never responded to a speeded trial (e.g. no keyboard).
n_no_response <- complete |> left_join(stroop_accuracy, by = "participant_id") |>
  filter(all_timed_out %in% TRUE) |> nrow()
if (n_no_response > 0) {
  message("Note: ", n_no_response,
          " completed session(s) had no response on any speeded trial. Check whether they ",
          "were taken on a device without a keyboard.")
}

if (nrow(keep) == 0) {
  message("No participants passed the exclusion criteria, so the analyses below will be ",
          "empty. With pilot data this usually just means too few participants so far.")
}
```

## Demographics

```{r demographics}
demo <- trials |>
  filter(task == "demographics", participant_id %in% keep$participant_id) |>
  mutate(resp = map(response, ~ fromJSON(.x))) |>
  transmute(
    participant_id,
    age = map_dbl(resp, ~ as.numeric(.x$age %||% NA)),
    gender = map_chr(resp, ~ as.character(.x$gender %||% NA)),
    native_english = map_lgl(resp, ~ isTRUE(.x$native_english))
  )

demo |> summarise(n = n(), mean_age = mean(age, na.rm = TRUE), sd_age = sd(age, na.rm = TRUE))
demo |> count(gender)
```

## Confirmatory analysis: Stroop effect

Key test (example; replace it with the one you preregistered): responses are slower on
incongruent than on congruent trials. For each participant, take the mean RT of correct test
trials in each condition, dropping anticipations under 200 ms (timeouts have no RT). Then
run a paired t-test on the per-participant difference, incongruent minus congruent.

```{r stroop}
rt_trials <- stroop_all |>
  filter(participant_id %in% keep$participant_id, correct, !is.na(rt), rt >= 200)
n_correct <- stroop_all |> filter(participant_id %in% keep$participant_id, correct) |> nrow()
message("RT analysis uses ", nrow(rt_trials), " of ", n_correct,
        " correct test trials (the rest were under 200 ms).")

# One row per participant. A participant with no usable trial in one condition gets NaN
# there and is dropped, rather than silently counted with half the data.
stroop_effect <- rt_trials |>
  group_by(participant_id) |>
  summarise(
    congruent = mean(rt[congruency == "congruent"]),
    incongruent = mean(rt[congruency == "incongruent"]),
    .groups = "drop"
  ) |>
  mutate(effect = incongruent - congruent) |>
  filter(!is.na(effect))

stroop_effect |>
  summarise(n = n(), mean_congruent = mean(congruent), mean_incongruent = mean(incongruent),
            mean_effect = mean(effect), sd_effect = sd(effect))

if (nrow(stroop_effect) >= 2) {
  tt <- t.test(stroop_effect$incongruent, stroop_effect$congruent, paired = TRUE)
  print(tt)
  # Effect size with a confidence interval, plus d_z (mean difference / SD of differences)
  # for comparison with the original study. Compare intervals, not just p values.
  tibble(
    effect_ms = unname(tt$estimate),
    ci_low = tt$conf.int[1],
    ci_high = tt$conf.int[2],
    d_z = mean(stroop_effect$effect) / sd(stroop_effect$effect)
  )
} else {
  message("Paired t-test not run: it needs at least two kept participants with usable ",
          "trials in both conditions (have ", nrow(stroop_effect), "). ",
          "With pilot data this is expected; come back once you have more data.")
}

if (nrow(stroop_effect) > 0) {
  stroop_effect |>
    pivot_longer(c(congruent, incongruent), names_to = "congruency", values_to = "mean_rt") |>
    ggplot(aes(x = congruency, y = mean_rt)) +
    geom_line(aes(group = participant_id), alpha = 0.3) +
    stat_summary(fun = mean, geom = "point", size = 3) +
    labs(x = "Congruency", y = "Mean correct RT (ms)",
         title = paste0("Stroop effect (n = ", nrow(stroop_effect), ")")) +
    theme_minimal()
} else {
  message("No Stroop RT data to plot yet.")
}
```

## Exploratory: accuracy and RT distributions

```{r exploratory}
kept_trials <- stroop_all |> filter(participant_id %in% keep$participant_id)

if (nrow(kept_trials) > 0) {
  print(kept_trials |>
          group_by(congruency) |>
          summarise(accuracy = mean(correct), timeouts = sum(timed_out), trials = n(),
                    .groups = "drop"))
  ggplot(rt_trials, aes(x = rt, fill = congruency)) +
    geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
    labs(x = "RT (ms), correct trials", fill = "Congruency") +
    theme_minimal()
} else {
  message("No test trials from kept participants yet.")
}
```

## Data quality

Client-side errors reported by the experiment page, if any.

```{r errors}
if (nrow(errors) > 0) {
  errors |> select(any_of(c("at", "uid", "trial_index", "message"))) |> head(20)
} else {
  message("No client-side errors were reported. This is what you want to see.")
}
```

## Session info

```{r session}
sessionInfo()
```
